---
title: "Titanic Analyses iter 1"
output: html_notebook
---

My first attempt at data science process
Identify objectives > Data collection and understanding > Data preparation >  Data Modeling > Evaluate Model > Conclusions and review

Identify Objecticces
1) Implment Data science flow on Titanic DS set to predict survival
2) analyze results and review what can be improved / what you did wrong

DATA COLLECTION AND UNDERSTANDING
Load Data

```{r}
train <- read.csv("train.csv", stringsAsFactors = F)
str(train)
test <- read.csv("test.csv", stringsAsFactors = F)
str(test)
```


Combine data for easier processing

```{r}
all <- rbind(train[names(train) != "Survived"], test)
str(all)
```

Lets visualize some of the data in particular Pclass, sex, and age factored by survival rate

```{r}
library(ggplot2)
library(gridExtra)

ggplot(train, aes(Pclass)) + geom_bar(position = "dodge", aes(fill = factor(train$Survived)))
ggplot(train, aes(Sex)) + geom_bar(position = "dodge", aes(fill = factor(train$Survived)))
ggplot(train, aes(Age)) + geom_histogram(position = "dodge", aes(fill = factor(train$Survived)))
```

DATA PREPARATION

Lets look for missing values using summary (this only takes care of Numericc Data)


```{r}
summary(all)
```
Age has 263 NA's, Fare has 1 NA. This takes care of numeric data.

For chr data lets sum up all "" values ("" seems to be the default string value for missing character values)

```{r}
Chr_cols <- c("Name", "Sex", "Ticket", "Cabin", "Embarked")
for(i in Chr_cols){
        
        print(paste(i, ": ", sum(all[i] == "")))
        
}
```

Embarked has 2 missing values and Cabin has 1014 missing values

The easiest values to replace are Embarked and Fare. We'll replace Embarked and Fare with the median for each

```{r}
all$Embarked[all$Embarked == ""] <- "S"
all$Fare[is.na(all$Fare)] <- 14.454
summary(all)

```

Now we want to attempt our first machine learning algorithm using the cleaned data.
I will be using Pclass, Sex, Embarked, and Age for my first attempt
First I need to make the data numeric / factors


```{r}
factor_col_names <- c("Pclass", "Sex", "Embarked")
for (i in factor_col_names){
        new_col_name <- paste("factor_", i)
        all[,new_col_name] <- factor(all[,i])
}
summary(all)
str(all)
```

DATA MODELING

Create Xtrain, Ytrain and Xtest in prepartion for glm

```{r}
train_names <- c("factor_ Pclass", "factor_ Sex")
Xtrain <- all[1:891, train_names]
Xtrain$Survived <- factor(train[,"Survived"])
Xtest <- all[892:1309, train_names]

model <- glm(Survived ~ ., data = Xtrain, family = binomial)
model_test <-factor(ifelse(predict(model, newdata = Xtrain[train_names]) > 0.5,1,0))
summary(model)
confusionMatrix(Xtrain$Survived, model_test)

compare <- data.frame(Model = model_test, Actual = factor(Xtrain$Survived))
head(compare)

```

CONCLUSIONS AND REVIEW

Even with only a few features we were able to achieve ~ 78% accuracy rate

Clearly we can use more features and there is more data to look at

Furthermore we did not create either a test/validation/test set so our data is most likeley over trained

we did not look at any regularization parameters or try to identify bias/variance

Other ML algorithms should also be considered (SVM, decision trees)
